\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{array}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{mdframed}
\usepackage{caption}
\usepackage{subcaption}

\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue}

\newcommand{\SQR}{\text{SQR}}
\newcommand{\CPSQR}{\text{CPSQR}}
\newcommand{\chiT}{\chi T / 2\pi}
\newcommand{\Fstrict}{F_{\text{strict}}}
\newcommand{\Fcpsqr}{F_{\text{CPSQR}}}
\newcommand{\nact}{N_{\text{active}}}

\definecolor{pass}{RGB}{0,130,0}
\definecolor{fail}{RGB}{180,0,0}
\definecolor{warn}{RGB}{200,120,0}

\newcommand{\passmark}{\textcolor{pass}{$\checkmark$}}
\newcommand{\failmark}{\textcolor{fail}{$\times$}}
\newcommand{\warnmark}{\textcolor{warn}{$\sim$}}

\title{\textbf{Unified Review: Multitone SQR Control and\\
Holographic Cluster-State Synthesis in Dispersive cQED}\\[0.5em]
\large A Critical Audit, Rerun, and Synthesis of Nine Prior Studies}

\author{Critical Review --- March 28, 2026}

\date{}

\begin{document}
\maketitle
\tableofcontents
\clearpage

%% ============================================================
\section{Overview and Scientific Questions}
%% ============================================================

This document audits, reruns, and unifies the nine most recent studies in this research
repository, all of which address one or both of the following related questions:

\begin{mdframed}
\textbf{Scientific Question 1 (SQR):}
Can a parameterized multitone qubit-drive waveform simultaneously realize
a \emph{strict ideal x-axis selective qubit rotation} (SQR)
$U_n = \exp(-i\theta_n X / 2)$
on each addressed Fock manifold $n \in \{0, 1, \ldots, \nact{-}1\}$
of a dispersive transmon-cavity system,
with no residual per-manifold conditional phase,
to a joint process fidelity exceeding 0.99?
And if so, under what conditions (pulse duration, number of addressed manifolds,
Hamiltonian model) does this succeed, and where does it break down?

\textbf{Scientific Question 2 (Cluster State):}
Can a structured native-gate sequence (D\,+\,R\,+\,SQR or D\,+\,R\,+\,CPSQR)
implement the restricted ground-sector holographic cluster-state transfer
$U_\text{joint}(\vert g\rangle \otimes \vert\psi\rangle) \approx \vert g\rangle \otimes H_c\vert\psi\rangle$
on $\{\vert g,0\rangle, \vert g,1\rangle\}$ to fidelity exceeding 0.99,
and how does the minimum circuit depth compare to the unconstrained GRAPE benchmark?
\end{mdframed}

The nine studies were carried out autonomously by a prior agent.
This review reexamines every study skeptically, checks all key claims against the
saved data and against rerun simulations, identifies weaknesses and inconsistencies,
and produces a unified set of evidence-backed conclusions.

%% ============================================================
\section{Physical Model and Shared Conventions}
%% ============================================================

\subsection{Hamiltonian}

All nine studies share the same dispersive transmon-cavity Hamiltonian:
\begin{equation}
  H = \omega_c a^\dagger a + \omega_q \vert e\rangle\langle e\vert
      + \frac{\alpha}{2}\vert e\rangle\langle e\vert(\vert e\rangle\langle e\vert - 1)
      - \chi a^\dagger a \vert e\rangle\langle e\vert
      - \frac{\chi'}{2} (a^\dagger a)^2 \vert e\rangle\langle e\vert
      + K (a^\dagger a)^2
\end{equation}
with parameters:
\begin{center}
\begin{tabular}{lll}
\toprule
Parameter & Symbol & Value \\
\midrule
Qubit frequency & $\omega_q / 2\pi$ & 6.150\,GHz \\
Cavity frequency & $\omega_c / 2\pi$ & 5.241\,GHz \\
Transmon anharmonicity & $\alpha / 2\pi$ & $-255$\,MHz \\
Dispersive shift & $\chi / 2\pi$ & $-2.84$\,MHz \\
Second-order dispersive shift & $\chi' / 2\pi$ & $-21$\,kHz \\
Cavity Kerr & $K$ & 0 (set to zero in all studies) \\
Transmon levels retained & $N_\text{tr}$ & 2 \\
Cavity levels & $N_\text{cav}$ & 4--12 (study dependent) \\
\bottomrule
\end{tabular}
\end{center}

\noindent The SQR studies use two model variants:
\texttt{chi\_only} (sets $\chi' = 0$) and \texttt{chi\_plus\_chiprime} (full model).
The cluster-state study uses the full model with an additional Kerr term $K/2\pi = -28$\,kHz
for the GRAPE benchmark only.

\subsection{SQR Target Definition}

A \textbf{strict ideal SQR} on $\nact$ addressed Fock manifolds requires:
\begin{equation}
  U_\text{joint}\big\vert_{n \in \text{addr}} = R_x(\theta_n)
  \equiv e^{-i\theta_n X / 2},
  \qquad
  U_\text{joint}\big\vert_{n \notin \text{addr}} = \mathbf{1}
\end{equation}
with \emph{identical rotation angle $\theta_n$ for all $n$} and \emph{zero residual Z phase}.
The minimum free parameter count for this target is $2\nact$ real numbers
(one $\theta_n$ and one phase-lock constraint per manifold).

A \textbf{relaxed CPSQR} (conditional phase SQR) allows arbitrary per-manifold
Z phases $\varphi_n$:
\begin{equation}
  U_\text{joint}\big\vert_{n} = e^{i\varphi_n Z/2} R_x(\theta_n)
\end{equation}
The CPSQR target is substantially easier because the optimizer is not required to
suppress conditional phase accumulation.

\subsection{Metric Hierarchy}

Multiple metrics appear across the nine studies, and their inconsistent use is
a major source of confusion.  In order of increasing stringency:

\begin{enumerate}
\item \textbf{Reduced/restricted process fidelity} --- fidelity of the per-manifold
  reduced effective-qubit unitary, ignoring cavity-mode correlations.
  \emph{Most optimistic; can be high even when the full joint operator is wrong.}
\item \textbf{Single-ground-state probe fidelity} --- state fidelity on
  $\vert g\rangle$ input only.  Misses all coherence errors.
\item \textbf{Spanning-quartet probe fidelity} --- mean state fidelity over
  $\{\vert g\rangle, \vert e\rangle, \vert{+x}\rangle, \vert{+y}\rangle\}$ inputs
  per manifold.  Sufficient to certify a per-manifold unitary up to global phase.
\item \textbf{Full joint process fidelity} --- process fidelity of the complete
  joint qubit-cavity unitary on all addressed and unaddressed manifolds simultaneously.
  \emph{Most stringent; the definitive metric.}
\end{enumerate}

\noindent A key finding from Study~4 (multi-input validation) is that
\emph{metrics 1 and 2 can yield high values while metric 4 is essentially zero}.
The definitive metric throughout this review is the full joint process fidelity.

%% ============================================================
\section{The Nine Studies: Goals, Methods, and Weaknesses}
%% ============================================================

\begin{center}
\begin{tabular}{rp{4cm}p{3.5cm}p{5cm}}
\toprule
\# & Study & Primary goal & Key weakness identified in this review \\
\midrule
1 & \texttt{corrected\_sqr\_conditioned\_rotation\_metric} &
  Establish correct qubit-effective metric for SQR optimization &
  Only thoroughly validates $\nact=1$; multi-manifold extension not demonstrated \\
\addlinespace
2 & \texttt{parameterized\_waveform\_residual\_z\_cancellation} &
  Test richer waveform families for Z-error suppression &
  Reports restricted process fidelity; multi-input validation (Study~4) shows
  the baseline result is essentially broken ($F_\text{joint} = 0.008$) \\
\addlinespace
3 & \texttt{ideal\_sqr\_direct\_vs\_echoed\_multitone} &
  First targeted strict ideal x-axis SQR; direct vs echoed &
  Best direct case has \texttt{optimizer\_success=False}; echoed does not demonstrably improve strict metric \\
\addlinespace
4 & \texttt{native\_multitone\_sqr\_fixed\_multiinput\_validation} &
  Validate saved artifacts from Studies 1--3 on multi-input probe ladder &
  Reveals that Studies 2 and 3 artifacts fail on superposition probes \\
\addlinespace
5 & \texttt{multitone\_sqr\_arbitrary\_fock\_conditional\_rotations} &
  Broader arbitrary SU(2) block targets; direct vs echoed &
  Mean process fidelity 0.35 (direct) and 0.22 (echoed); echoed is \emph{worse} on average \\
\addlinespace
6 & \texttt{strong\_validation\_arbitrary\_fock\_conditional\_rotations} &
  5 waveform families; strict and relaxed CPSQR evaluation &
  Higher-expressivity benchmark yields zero gain; reason not investigated \\
\addlinespace
7 & \texttt{native\_rich\_multitone\_sqr\_cpsqr\_feasibility} &
  Most comprehensive: 10 families, 3-stage screen, both models &
  High quartet validation on separate manifolds does not certify joint operator
  (one case: quartet $= 0.9999997$, joint $= 0.868$) \\
\addlinespace
8 & \texttt{the\_definitive\_ideal\_sqr\_gate\_study} &
  Aggregate and synthesize all prior SQR results &
  Aggregation-only pass; no new optimizations run; practical ranking
  uses decoherence proxy, not actual drift sweeps \\
\addlinespace
9 & \texttt{cluster\_state\_holographic\_unified} &
  Holographic cluster-state ground-sector synthesis &
  Old full-target winners completely fail for ground-sector task (ancilla excitation $\approx 1$);
  no open-system validation for ground-sector branch yet \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Study 1: Corrected Metric (\texttt{corrected\_sqr\_conditioned\_rotation\_metric})}

\textbf{Goal.} Establish that the correct objective for SQR optimization is the
reduced effective-unitary weighted process fidelity, not raw state fidelity.
Verify the phase convention used by \texttt{cqed\_sim}'s multitone ansatz
(specifically, confirm that $\phi$ is an axis-angle, not a quadrature offset)
and demonstrate analytic and gradient-based correction methods.

\textbf{Methods.} A 4-level target profile with varying $\theta_n$ and $\phi_n$ per manifold.
Four optimization methods compared: baseline (no correction), analytic warm-start,
state-based optimization, and unitary-based optimization.

\textbf{Results.}  For $\nact = 1$, the unitary-optimized case achieves
$F_\text{joint} = 0.9999999$ and zero residual $Z$ or transverse error.
For $\nact = 4$, the baseline fidelity is poor but the corrected metric guides
the optimizer to better solutions.

\textbf{Weakness.}  The study validates the metric and convention definitively for
$\nact = 1$ but the multi-manifold cases ($\nact = 4$) are described but not
independently validated.  The study summary shows per-manifold convergence checks
but not the full joint process fidelity at $\nact = 4$.
Study~4 multi-input validation confirms that the corrected-metric
unitary-optimized case achieves 1.0000 spanning-quartet fidelity,
which is the strongest per-manifold validation available short of a full GRAPE comparison.

\subsection{Study 2: Parameterized Waveform (\texttt{parameterized\_waveform\_residual\_z\_cancellation})}

\textbf{Goal.}  Pilot study on five richer waveform families
(baseline multitone, symmetric two-segment, echoed, complex-envelope, basis-expanded)
to suppress residual-Z errors while targeting family-C (structured) and
family-D (random) block-diagonal conditional rotations.

\textbf{Results.}  Best result: baseline multitone on family-C, $\nact = 2$, $\chiT = 3$,
restricted process fidelity $= 0.842$, average gate fidelity $= 0.873$.

\textbf{Critical weakness (confirmed by Study~4 rerun).}
The baseline multitone result on family-D ($\nact = 4$) achieves
restricted process fidelity $= 0.008$ (essentially zero).
Yet the single-ground-state probe fidelity is 0.574 and the spanning-quartet mean
drops to 0.443.
\emph{This study's headline metric is the restricted process fidelity, which is
completely misleading for the hard targets.  The waveform is not implementing
the intended operation.}  Richer waveforms did not fix this; the study
correctly notes that family-D targets are hard but does not escalate the failure
to a primary conclusion.

\subsection{Study 3: Ideal SQR Direct vs Echoed (\texttt{ideal\_sqr\_direct\_vs\_echoed\_multitone})}

\textbf{Goal.}  First study specifically targeting strict ideal x-axis SQR
(all $\theta_n$ equal, all $\phi_n = 0$).  Audit finding: earlier studies used
arbitrary SU(2) block targets, not the strict x-axis target.

\textbf{Results.}
\begin{itemize}
\item Best direct: average gate fidelity 0.7245 on $\nact = 2$, $\chiT = 5.0$
  (chi\_only model).  Per-block fidelities are highly uneven:
  block 0 process fidelity 0.794, block 1 only 0.531.
  \textbf{Optimizer success flag: False.}  The solution is at a poor local minimum.
\item Echoed constructions: do not materially improve strict SQR fidelity.
  The mean strict process fidelity for echoed is lower than for direct on most cases.
\item Study~4 multi-input validation for this study's best artifact
  gives joint process fidelity $= 0.824$, spanning-quartet mean $= 0.875$.
  The improvement over Study~2 is real but still far from 0.99.
\end{itemize}

\textbf{Weakness.}  The convergence failure is acknowledged but not resolved.
The study presents echoed results as comparable to direct without emphasizing that
echoed constructions fail for strict SQR because the inserted $X_\pi$ pulse
is strongly manifold-dependent (see Study~8 diagnostics below).

\subsection{Study 4: Multi-Input Validation (\texttt{native\_multitone\_sqr\_fixed\_multiinput\_validation})}

\textbf{Goal.}  Validate saved artifacts from Studies 1--3 using a three-tier
probe ladder: single ground, selected pair (g, +x), and spanning quartet.

\textbf{Results.}  \textbf{This is the most critical audit study in the collection.}
Confirmed results from the rerun of \texttt{run\_study.py} on 2026-03-28:

\begin{center}
\begin{tabular}{llllll}
\toprule
Source study & Construction & $F_\text{joint}$ & Quartet mean & Quartet min & Assessment \\
\midrule
Study 2 & baseline multitone & 0.0084 & 0.443 & 0.003 & \failmark{} fails \\
Study 5 & direct multitone & 0.601 & 0.734 & 0.346 & \warnmark{} partial \\
Study 3 & ideal SQR direct & 0.824 & 0.875 & 0.720 & \warnmark{} partial \\
Study 1 & corrected unitary-opt & 0.9986 & 1.000 & 1.000 & \passmark{} passes \\
\bottomrule
\end{tabular}
\end{center}

The Study~2 case is especially striking: restricted process fidelity was reported
as 0.008 (correctly low), but the single-ground-state fidelity was 0.574,
creating the appearance of partial success that is not real.
The unitary-optimized corrected-metric case (Study~1) is the only one that
passes the multi-input validation to high fidelity.

\subsection{Study 5: Arbitrary Fock-Conditional Rotations (\texttt{multitone\_sqr\_arbitrary\_fock\_conditional\_rotations})}

\textbf{Goal.}  Broader study with arbitrary SU(2) block targets (not just
strict ideal SQR), covering $\nact \in \{2,3,4,5\}$, durations
$\chiT \in \{1,3,5,7\}$, and both chi-only and chi+$\chi'$ models.
Single-pulse versus explicit echoed constructions.

\textbf{Results.}  Mean process fidelity: 0.350 for single-pulse, 0.225 for echoed.
Best single-pulse: family-A, $\nact = 2$, chi\_only, $F_\text{avg} = 0.714$.

\textbf{Weakness.}  The conclusion that ``single pulse remains best overall''
is correct in mean terms but undersells the depth of the failure.
A mean of 0.35 represents mostly failed optimizations.
The result that echoed is \emph{worse} on average is important but not given
a satisfying explanation; the study attributes it to waveform degrees of freedom
without investigating whether the $X_\pi$ quality is the primary cause.

\subsection{Study 6: Strong Validation (\texttt{strong\_validation\_arbitrary\_fock\_conditional\_rotations})}

\textbf{Goal.}  Five waveform families (single-pulse Gaussian, independent-half echo,
symmetry-constrained echo, segmented relaxed CPSQR, basis-expanded benchmark).
114 evaluations. Strict and relaxed objectives evaluated jointly.

\textbf{Key results.}
\begin{itemize}
\item Best strict case: single-pulse Gaussian, $F_\text{strict} = 0.9957$,
  $\nact = 2$, $\chiT = 5.0$, chi\_plus\_chiprime, in-plane-axes target.
\item Segmented relaxed CPSQR: achieves joint process fidelity $\approx 1.000$
  on hard targets ($\nact = 3$, random).
\item Higher-expressivity basis-expanded benchmark: \textbf{zero improvement over baseline}.
\end{itemize}

\textbf{Weakness.}  The zero gain from the basis-expanded family is suspicious
and not investigated.  It suggests either a broken optimizer path, an
insufficient budget, or a genuine ansatz-vs-physics mismatch, none of which
are resolved.  The ``best strict case'' at 0.9957 is the headline result of this study,
but it comes from the restricted view only and is not independently multi-input-validated
within this study's own framework.

\subsection{Study 7: Native/Rich Multitone Feasibility (\texttt{native\_rich\_multitone\_sqr\_cpsqr\_feasibility})}

\textbf{Goal.}  Most comprehensive SQR study.  Screens 10 waveform families in three stages:
initial screen on chi+$\chi'$ model, comparison across both models and up to $\nact = 4$,
and duration refinement sweep.  Introduces the full joint process fidelity
and the negative control recomputation.

\textbf{Key findings (confirmed against saved data).}
\begin{itemize}
\item \textbf{Best strict-SQR}: reduced\_unitary\_direct and native\_direct\_strict,
  $F_\text{strict,joint} = 0.9990$ on chi\_only, $\nact = 2$, $\chiT = 6.0$,
  smooth\_x target.
  On $\chiT = 5.0$: $F_\text{strict,joint} = 0.9988$.
  Both chi-only and chi+$\chi'$ models give the same best result.
\item \textbf{Best CPSQR}: echoed\_independent,
  $F_\text{CPSQR,joint} = 1.0000$ on chi\_plus\_chiprime, $\nact = 2$,
  $\chiT = 5.0$, staggered\_x target.
  Also achieves 1.0000 for chi\_only and for $\nact = 3$.
\item \textbf{Critical caveat}: On one case (echoed\_asymmetric,
  chi\_only, $\nact = 2$, staggered\_x), the reduced quartet fidelity is 0.9999997
  but the full joint process fidelity is only 0.8677.
  \emph{The reduced-manifold view is not a safe proxy for joint fidelity.}
\item \textbf{Performance scaling with $\nact$}:
  $\nact = 2$: best $= 0.9990$, mean $= 0.721$;
  $\nact = 3$: best $= 0.9975$, mean $= 0.675$;
  $\nact = 4$: best $= 0.9948$, mean $= 0.646$.
\item Echoed families have \emph{lower} mean strict fidelity than direct families
  for all $\nact$, confirming Study~5.
\item The negative control recomputation showed that the old restricted metric
  (0.4949) underestimated performance under the patched cqed\_sim
  (correct quartet = 0.875); metric inconsistency between studies is real.
\end{itemize}

\textbf{Weakness.}  The study does not include a fresh GRAPE or unconstrained
time-domain comparison on the hard cases, so it cannot distinguish
ansatz-limited failure from physics-limited failure for the n\_active $\ge 3$
cases.

\subsection{Study 8: Definitive Aggregation (\texttt{the\_definitive\_ideal\_sqr\_gate\_study})}

\textbf{Goal.}  Load and synthesize all four key source studies into a single
machine-readable table. Add standalone diagnostics: X$_\pi$ pulse characterization,
spectral crowding analysis, parameter scaling, addressing window scaling fits.

\textbf{Rerun (2026-03-28).}  The full rebuild completed in 36.98\,s (wall clock).
All four sanity checks pass:
\begin{itemize}
\item Baseline best direct: $F_\text{strict} = 0.6557$ ($\approx 0.65$ as expected)
\item Native-rich strict $> 0.99$: value $= 0.9990$ \passmark
\item Native-rich CPSQR $> 0.999$: value $= 1.0000$ \passmark
\item Row counts: 64\,(Study~3) + 270\,(Study~5) + 348\,(Study~7) + 240\,(Study~2) = 922 total \passmark
\end{itemize}

\textbf{Weakness.}  The study is an aggregation pass only.  No new optimizations were
run.  The ``practical ranking'' uses a decoherence-proxy score, not actual drift sweeps.
The IMPROVEMENTS.md lists three P1 gaps: no full echoed-grid rerun with the
compromise $X_\pi$ pulse, no fresh drift analysis, and no GRAPE/unconstrained
upper bound.

\textbf{New diagnostics produced.}
\begin{enumerate}
\item \textbf{$X_\pi$ characterization}: The refocusing pulse (40\,ns Gaussian,
  baseline parameters) degrades strongly with Fock level:
  level 0: $F = 0.9993$; level 1: $F = 0.9716$; level 2: $F = 0.8926$;
  level 3: $F = 0.7734$; level 4: $F = 0.6298$.
  A compromise pulse (shorter duration, adjusted sigma) improves the
  worst-manifold fidelity to $0.919$ but does not certify echoed strict-SQR.
\item \textbf{Spectral crowding}: At $\chiT = 3.0$, the tone bandwidth equals
  twice the manifold spacing ($\text{crowding ratio} = 2.0$) for all levels.
  This is a constant spectral bottleneck that does not worsen with manifold index.
\item \textbf{Parameter count}: Minimum free parameters for ideal SQR = $2\nact$.
  The current families use $6\,\text{--}\,18$ parameters for $\nact = 2$,
  giving overcomplete representations that the optimizer handles well.
\end{enumerate}

\subsection{Study 9: Holographic Cluster-State Synthesis (\texttt{cluster\_state\_holographic\_unified})}

\textbf{Goal.}  Implement two control problems using structured gate families
D\,+\,R\,+\,SQR and D\,+\,R\,+\,CPSQR:
(a) the full joint logical transfer unitary SWAP\,$\cdot$\,CZ\,$\cdot$\,(H$\otimes$I)
on $\{\vert g,0\rangle, \vert g,1\rangle, \vert e,0\rangle, \vert e,1\rangle\}$; and
(b) the restricted ground-sector transfer
$U_\text{joint}(\vert g\rangle \otimes \vert\psi\rangle) \approx \vert g\rangle \otimes H_c\vert\psi\rangle$
on $\{\vert g,0\rangle, \vert g,1\rangle\}$.

\textbf{Critical finding.}  The full-target winners discovered in objective~(a)
\emph{completely fail} on objective~(b): their ground-sector fidelities collapse to
$\approx 0.50$ with ancilla excitation near 1.  This is because those solutions
route amplitude through the excited ancilla sector.  The two objectives are
genuinely different control problems.

\textbf{Key ground-sector results (from saved data, confirmed consistent):}

\begin{center}
\begin{tabular}{llllll}
\toprule
Family & Blocks & Ordering & $F_\text{gnd,12}$ & Ancilla$_\text{worst}$ & Leakage$_\text{worst}$ \\
\midrule
D+R+SQR & 2 & SRD & 0.9876 & $8.9\times10^{-4}$ & $2.97\times10^{-2}$ \\
D+R+SQR & 3 & RDS & 0.9997 & $4.7\times10^{-4}$ & $5.97\times10^{-4}$ \\
D+R+SQR & 4 & RDS & 0.9965 & $3.75\times10^{-3}$ & $6.96\times10^{-3}$ \\
D+R+SQR & 5 & RDS & 0.9999 & $4.1\times10^{-5}$ & $1.77\times10^{-4}$ \\
\midrule
D+R+CPSQR & 1 & CPRD & 0.8615 & $\approx 0$ & $2.59\times10^{-1}$ \\
D+R+CPSQR & 2 & CPRD & 0.9887 & $1.6\times10^{-9}$ & $3.05\times10^{-2}$ \\
D+R+CPSQR & 3 & DRCP & 0.9979 & $1.6\times10^{-3}$ & $4.20\times10^{-3}$ \\
D+R+CPSQR & 4 & RDCP & 0.9999 & $9.0\times10^{-5}$ & $1.76\times10^{-4}$ \\
D+R+CPSQR & 5 & DRCP & 0.9995 & $8.99\times10^{-4}$ & $1.12\times10^{-3}$ \\
\bottomrule
\end{tabular}
\end{center}

Both families first cross the 0.99 threshold at 3 blocks.
The 2-block frontier (SQR: 0.988, CPSQR: 0.989) is just below threshold.

The GRAPE benchmark at $N_\text{cav} = 12$ for the \emph{full target} reaches
replay fidelity 0.976 at 800\,ns.  No like-for-like GRAPE baseline was run
for the restricted ground-sector objective.

\textbf{Weaknesses.}
\begin{itemize}
\item The ground-sector follow-up used a warm-start from the best full-target solutions,
  not a fresh unrestricted search.  A dedicated low-depth search could change the
  minimum-depth conclusion.
\item No open-system validation for any of the ground-sector winners.
\item The GRAPE frontier is a full-target benchmark only; comparing structured-family
  ground-sector fidelity to full-target GRAPE fidelity is not a like-for-like comparison.
\item The non-monotonic depth dependence (4-block SQR $= 0.997 <$ 3-block SQR $= 0.9997$)
  reflects optimization variance, not a fundamental non-monotonicity.
\end{itemize}

%% ============================================================
\section{Unified Methodology}
%% ============================================================

\subsection{Notation and Target Conventions (Unified)}

Throughout this review, all fidelity claims use the \textbf{full joint process fidelity}
$F_\text{strict,joint}$ for strict SQR and $F_\text{CPSQR,joint}$ for relaxed CPSQR,
defined as:
\begin{equation}
  F_\text{joint} = \frac{1}{d^2} \left\vert \text{tr}\left(U_\text{target}^\dagger
  U_\text{optimized}\right)\right\vert^2
\end{equation}
where the trace is over all $d$ joint qubit-cavity logical levels.
Reduced metrics (qubit-only, per-manifold, single-state) are shown as supplementary
information only.

The targeted SQR target is:
\begin{equation}
  U_\text{target,strict} = \bigotimes_{n=0}^{\nact-1} R_x(\pi/2) \otimes \mathbf{1}_\text{unaddressed}
\end{equation}
(uniform $\pi/2$ rotation on all addressed manifolds for ``smooth\_x'' targets;
staggered angles for ``staggered\_x'').

\subsection{Waveform Families Compared}

The following families appear across the studies.
This table unifies naming conventions:

\begin{center}
\begin{tabular}{lll}
\toprule
Internal family name & Display name & Parameters ($\nact = 2$) \\
\midrule
\texttt{gaussian\_seed} & direct seed & 12 \\
\texttt{native\_direct\_strict} & native direct & 12 \\
\texttt{reduced\_unitary\_direct} & reduced unitary & 12 \\
\texttt{symmetric\_two\_segment} & two-segment direct & 18 \\
\texttt{complex\_envelope} & rich envelope & 15 \\
\texttt{basis\_expanded} & basis-expanded & 15 \\
\texttt{echoed\_independent} & independent echo & 24 \\
\texttt{echoed\_asymmetric} & asymmetric echo & 24 \\
\texttt{echoed\_symmetric} & symmetric echo & 12 \\
\texttt{echoed\_cpsqr} & CPSQR echo & 24 \\
\texttt{segmented\_relaxed} & segmented CPSQR & variable \\
\bottomrule
\end{tabular}
\end{center}

%% ============================================================
\section{Rerun Results and Reproducibility}
%% ============================================================

\subsection{What Was Rerun}

\begin{enumerate}
\item \textbf{Definitive aggregation study} (\texttt{run\_full\_study.py --full}):
  Completed in 36.98\,s on 2026-03-28.
  All sanity checks pass.  Results match saved data exactly.
  \textbf{Reproducible.} \passmark

\item \textbf{Multi-input validation study} (\texttt{run\_study.py}):
  Completed on 2026-03-28.
  Results match saved data exactly (see Table in Section~3.4).
  \textbf{Reproducible.} \passmark

\item \textbf{Native-rich key metrics} (read from saved JSON, spot-checked via manual queries):
  Best strict $= 0.9990$, best CPSQR $= 1.0000$, performance scaling with $\nact$ confirmed.
  \textbf{Consistent with claims.} \passmark

\item \textbf{Cluster-state ground-sector summary} (read from saved JSON):
  Both families cross 0.99 at 3 blocks.  Best absolute: CPSQR 4-block $= 0.9999$.
  \textbf{Consistent with claims.} \passmark
\end{enumerate}

\subsection{What Was Not Rerun}

\begin{itemize}
\item The full native-rich 3-stage optimization pipeline
  (\texttt{run\_study.py} for \texttt{native\_rich}): computationally expensive,
  saved results are internally consistent, not rerun.
\item The cluster-state design-space search (8-worker multiprocessing run):
  saved results are self-consistent, not rerun.
\item The GRAPE frontier extension: computationally expensive, saved results used.
\item Any fresh unconstrained optimization or GRAPE run for strict SQR
  (identified as the critical missing experiment in all IMPROVEMENTS.md files).
\end{itemize}

\subsection{Metric Inconsistency Across Studies}

A notable finding is that the ``restricted process fidelity'' computed in Studies~2--5
is systematically more optimistic than the full joint process fidelity computed in
Studies~7--8.  For the negative-control case (chi+$\chi'$, $\nact = 3$, $\chiT = 3.0$,
direct multitone):
\begin{itemize}
\item Reported restricted process fidelity (Study~3): $F = 0.4949$
\item Recomputed joint process fidelity (Study~7): $F = 0.8242$
\item Recomputed spanning-quartet mean (Study~7): $F = 0.875$
\end{itemize}
The direction of the discrepancy is counterintuitive: the full-system metric
is \emph{higher} than the old restricted metric.  This reflects a cqed\_sim
API patch that corrected a phase convention between the two eras.
\textbf{No quantitative result from Studies~2--5 can be directly compared to
Studies~7--8 without metric normalization.}

%% ============================================================
\section{Key Quantitative Results}
%% ============================================================

\subsection{Best Strict Ideal-SQR Results}

\begin{center}
\begin{tabular}{llllll}
\toprule
Family & Model & $\nact$ & $\chiT$ & $F_\text{strict,joint}$ & $F_\text{CPSQR,joint}$ \\
\midrule
reduced\_unitary\_direct & chi\_only & 2 & 6.0 & \textbf{0.9990} & 0.9997 \\
native\_direct\_strict & chi\_only & 2 & 6.0 & \textbf{0.9990} & 0.9997 \\
reduced\_unitary\_direct & chi+$\chi'$ & 2 & 6.0 & \textbf{0.9990} & 0.9997 \\
native\_direct\_strict & chi+$\chi'$ & 2 & 6.0 & \textbf{0.9990} & 0.9997 \\
gaussian\_seed & chi+$\chi'$ & 2 & 5.0 & 0.9985 & 0.9999 \\
reduced\_unitary\_direct & chi\_only & 3 & 5.0 & 0.9975 & 0.9994 \\
reduced\_unitary\_direct & chi\_only & 4 & (long) & 0.9948 & --- \\
\midrule
echoed\_independent & chi\_only & 2 & 3.0 & 0.8671 & \textbf{1.0000} \\
echoed\_independent & chi+$\chi'$ & 2 & 5.0 & 0.8677 & \textbf{1.0000} \\
echoed\_independent & chi\_only & 3 & 5.0 & 0.6784 & \textbf{1.0000} \\
echoed\_asymmetric & chi\_only & 3 & 3.0 & 0.6769 & \textbf{1.0000} \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Main observation:}  For strict SQR, direct families (reduced\_unitary\_direct,
native\_direct\_strict) dominate at $\nact = 2$--$4$.
Echoed families dominate for CPSQR but are worse for strict SQR because
the inserted $X_\pi$ is manifold-dependent.

\subsection{Performance Scaling with Number of Addressed Manifolds}

From the all\_results.json of Study~7 (432 rows, all families, all $\chiT$):

\begin{center}
\begin{tabular}{lllll}
\toprule
$\nact$ & Best strict (any fam, any $\chiT$) & Mean strict & Best CPSQR \\
\midrule
2 & 0.9990 & 0.721 & 1.000 \\
3 & 0.9975 & 0.675 & 1.000 \\
4 & 0.9948 & 0.646 & 1.000 \\
\bottomrule
\end{tabular}
\end{center}

The \emph{best} strict fidelity remains above 0.99 for all tested $\nact$,
but only at long durations ($\chiT \ge 3$--$5$) and only for specific
(smooth, staggered) target families.
The \emph{mean} strict fidelity across all cases is 0.65--0.72,
indicating that the 0.99 regime is narrow and not the typical outcome.

\subsection{Minimum Duration for High Fidelity (Direct Families)}

For native\_direct\_strict, chi\_only model, smooth\_x target:

\begin{center}
\begin{tabular}{lll}
\toprule
$\nact$ & Threshold & Minimum $\chiT / 2\pi$ achieving threshold \\
\midrule
2 & 0.99 & 3.0 \\
2 & 0.995 & 3.0 \\
2 (CPSQR) & 0.99 & 2.0 \\
\bottomrule
\end{tabular}
\end{center}

At $|\chi|/2\pi = 2.84$\,MHz, $\chiT = 3$ corresponds to a pulse duration of
$T = 3 / (2\pi \times 2.84 \times 10^6\,\text{Hz}) \approx 168$\,ns.

\subsection{X$_\pi$ Pulse Quality Across Manifolds}

The 40\,ns Gaussian $X_\pi$ pulse used in echoed constructions
has sharply degrading quality with Fock level:

\begin{center}
\begin{tabular}{lll}
\toprule
Fock level $n$ & Process fidelity (baseline) & Process fidelity (robust 20\,ns) \\
\midrule
0 & 0.9993 & 0.9996 \\
1 & 0.9716 & 0.9808 \\
2 & 0.8926 & 0.9325 \\
3 & 0.7734 & 0.8788 \\
4 & 0.6298 & 0.7875 \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{This is the primary reason echoed constructions fail for strict SQR
when $\nact \ge 2$.}
Even the compromise pulse only improves the worst-manifold $X_\pi$ quality to $\approx 0.919$.
A full echoed strict-SQR success would require a manifold-specific $X_\pi$ sequence,
which is itself a selective rotation problem.

%% ============================================================
\section{Synthesis: What the Program Tells Us}
%% ============================================================

\subsection{The Core Physics Picture}

The dispersive Hamiltonian encodes a per-manifold shift
$\delta_n = -\chi n - \frac{\chi'}{2}n(n-1)$ on the qubit transition frequency.
An ideal strict SQR at duration $T$ requires the qubit to undergo a uniform
rotation $\theta_n = \theta_0$ for all addressed $n$ while accumulating zero
net conditional phase.  The phase accumulation rate is $\dot\phi_n = \delta_n$,
which grows linearly with $n$.

A multitone ansatz addresses this by placing one drive tone per addressed manifold.
The minimum condition for zero conditional phase is that each tone produces the correct
transverse rotation while producing zero longitudinal (Z) rotation on its own manifold
and zero cross-talk on other manifolds.  This is achievable in principle at long enough
durations (when spectral crowding is not limiting), which explains the experimental observation
that $F_\text{strict} > 0.99$ is achievable at $\chiT \ge 3$ for $\nact = 2$.

For larger $\nact$, the control problem becomes harder because more manifolds must be
simultaneously addressed with narrow tones, and the spectral crowding ratio (always
$\approx 2.0$ for uniform $|\chi|$ spacing) does not improve with $\nact$.

\subsection{Why Echoed Constructions Fail for Strict SQR}

The toggle-frame argument for echoed cancellation of conditional phase
assumes that the inserted $X_\pi$ pulse is a perfect bit-flip on every
addressed manifold.  Because the dispersive Hamiltonian is manifold-dependent,
no single-frequency $X_\pi$ pulse can be perfect on all manifolds simultaneously.
The $X_\pi$ pulse quality degrades from 0.999 at level 0 to 0.630 at level 4
(see above), which means the toggling-frame cancellation argument breaks down
precisely where it is most needed (higher photon numbers).

This finding, first identified clearly in Study~8 and now supported by the
standalone $X_\pi$ characterization, explains why:
\begin{enumerate}
\item Echoed constructions have \emph{lower mean strict fidelity} than direct
  constructions (Studies~5, 6, 7, confirmed).
\item Echoed constructions have \emph{near-perfect CPSQR fidelity}: the CPSQR
  target explicitly allows the residual Z phases that the imperfect $X_\pi$
  fails to cancel.
\item There is an inherent tension: achieving strict SQR via echoing requires a
  manifold-selective $X_\pi$, which is itself a harder form of SQR.
\end{enumerate}

\subsection{Relation to the Cluster-State Problem}

The cluster-state study is partially decoupled from the SQR program:
it uses SQR and CPSQR as primitive building blocks but optimizes the
full structured sequence rather than characterizing the primitives in isolation.
The key result --- that both families first cross 0.99 at 3 blocks for the
restricted ground-sector task --- is obtained by optimizing the full sequence,
not by composing pre-optimized SQR blocks.

The poor quality of the full-target solutions for the ground-sector task
(ancilla excitation $\approx 1$) is a structural finding that holds regardless
of how well the individual SQR primitives work.  It shows that the two
objectives (full transfer vs ground-sector transfer) are genuinely orthogonal.

%% ============================================================
\section{Final Conclusions}
%% ============================================================

\subsection{Claims That Are Robustly Supported}

\begin{enumerate}
\item \textbf{Strict ideal x-axis SQR at $\nact = 2$ with $\chiT \ge 3$:}
  Both the chi\_only and chi+$\chi'$ dispersive models admit strict ideal SQR
  solutions above $F = 0.99$ using the reduced\_unitary\_direct and
  native\_direct\_strict waveform families.
  Best observed: $F_\text{strict,joint} = 0.9990$ at $\chiT = 6.0$, $\nact = 2$.
  This result was confirmed from the saved data and is consistent with the
  definitive study rebuild.

\item \textbf{Strict SQR above 0.99 requires long enough pulses:}
  At $\nact = 2$, $F > 0.99$ requires $\chiT \ge 3.0$.  Shorter pulses fall below
  this threshold.

\item \textbf{Relaxed CPSQR achieves near-unit fidelity for all tested $\nact$:}
  The echoed\_independent family achieves $F_\text{CPSQR,joint} = 1.0000$ for
  $\nact = 2$--$3$ at $\chiT \ge 3$, on staggered\_x targets.
  This is confirmed by Study~7 and is consistent with all related studies.

\item \textbf{Echoed constructions fail for strict SQR due to $X_\pi$ manifold dependence:}
  The $X_\pi$ refocusing pulse fails increasingly at higher Fock levels.
  This is the primary reason echoed constructions have lower strict fidelity
  than direct constructions.  This finding is supported by both the
  averaged statistics and the explicit $X_\pi$ characterization.

\item \textbf{Multi-input validation is required to certify a gate:}
  Study~4 confirms that solutions passing single-state probes or restricted
  process fidelity can fail catastrophically on superposition probes.
  The reduced/restricted metric can be misleading in either direction
  (too optimistic or, with a convention patch, too pessimistic).

\item \textbf{Cluster state ground-sector task requires $\ge 3$ blocks:}
  Both D+R+SQR and D+R+CPSQR families first exceed $F = 0.99$ at 3 blocks
  under the restricted ground-sector objective.
  Best absolute fidelity: D+R+CPSQR at 4 blocks, $F_\text{gnd,12} = 0.9999$.

\item \textbf{Full-target structured solutions completely fail the ground-sector task:}
  Old full-target winners have ancilla excitation near 1 under the ground-sector
  objective.  The two problems are genuinely different.
\end{enumerate}

\subsection{Claims That Are Uncertain}

\begin{enumerate}
\item \textbf{Performance at $\nact \ge 5$:}
  No study tested $\nact = 5$ with the joint metric.  Study~5 included
  $\nact = 5$ but used an unreliable restricted metric.  Whether strict SQR
  above 0.99 is achievable at $\nact = 5$ is unknown.

\item \textbf{Whether strict SQR failures are ansatz-limited or physics-limited:}
  At $\nact = 3$--$4$, the best strict results are 0.9948--0.9975 and only
  at specific (long) durations.  Without a GRAPE or unconstrained upper bound,
  it is unknown whether the current direct families are near the physical limit
  or whether a better ansatz (e.g., GRAPE-optimized or time-domain free) would
  achieve significantly higher fidelity.

\item \textbf{Robustness to parameter drift:}
  The best solutions are numerically clean but have not been tested against
  systematic drift in $\chi$, $\omega_q$, or drive amplitude.
  The IMPROVEMENTS.md files list drift analysis as a P1 gap for Studies 7 and 8.

\item \textbf{Hardware relevance of the cluster-state result:}
  The cluster-state ground-sector winners have excellent closed-system fidelity
  but have not been validated with open-system (realistic decoherence) models.
  The GRAPE frontier open-system fidelity at 800\,ns is 0.919, substantially
  below the closed-system replay of 0.976.

\item \textbf{Minimum depth for the 2-block cluster-state frontier:}
  Both families are just below 0.99 at 2 blocks (SQR: 0.988, CPSQR: 0.989).
  Whether a fresh low-depth search could cross 0.99 at 2 blocks is unresolved.
\end{enumerate}

\subsection{Claims That Are Not Supported}

\begin{enumerate}
\item \textbf{``Strict SQR is generally achievable across the tested grid.''}
  The mean strict process fidelity across all tested cases is 0.65--0.72,
  with the 0.99+ regime confined to $\nact = 2$ at long durations.
  This is not a general success.

\item \textbf{``Richer waveform families improve strict SQR over the baseline.''}
  The complex-envelope and basis-expanded families do not improve over the
  native direct family for strict SQR (Study~7 comparison stage).
  The basis-expanded family provided zero gain even in Study~6.

\item \textbf{``Echoed constructions achieve strict SQR.''}
  Echoed families have \emph{lower} mean strict fidelity than direct families
  for all tested conditions.  While echoed constructions reliably achieve CPSQR,
  they do not achieve strict SQR.

\item \textbf{Any robustness claims from the ``practical ranking''} in the definitive study.
  That ranking uses a decoherence-proxy score only and has not been validated
  against actual noise or drift.

\item \textbf{``The 2-block cluster-state boundary can be crossed with existing families.''}
  Current data show 2-block performance below 0.99 for both families.
  The IMPROVEMENTS.md explicitly calls for a fresh low-depth search rather than
  concluding that 3 blocks is the fundamental minimum.
\end{enumerate}

%% ============================================================
\section{Recommended Next Steps}
%% ============================================================

In priority order:

\begin{enumerate}
\item \textbf{[P1] Run a GRAPE/unconstrained upper bound for strict SQR at
  $\nact = 3$--$4$, $\chiT = 3$--$5$.}
  This is the most critical missing experiment.
  Until an unconstrained comparison is available, the question of whether current
  failures are ansatz-limited or physics-limited cannot be answered.

\item \textbf{[P1] Perform parameter-drift sensitivity analysis on the best
  direct strict-SQR solutions.}
  Specifically: sweep $\chi \pm 5\%$, $\omega_q \pm 1$\,MHz, drive amplitude
  $\pm 5\%$ around the best $\nact = 2$ and $\nact = 3$ solutions.

\item \textbf{[P2] Run the full echoed-grid rerun with the compromise $X_\pi$ pulse.}
  The standalone $X_\pi$ characterization (Study~8) identified a shorter robust
  pulse with better manifold uniformity.  Running the echoed constructions with
  this pulse is a well-defined next experiment.

\item \textbf{[P2] Run a fresh 2-block low-depth cluster-state search}
  directly for the ground-sector objective (not warm-started from full-target solutions).

\item \textbf{[P2] Open-system validation for the 3-block and best-absolute
  cluster-state ground-sector winners.}
  The closed-system $F > 0.999$ does not guarantee good hardware performance.

\item \textbf{[P3] Extend the SQR study to $N_\text{tr} = 3$ transmon levels.}
  All current studies use $N_\text{tr} = 2$.  The second excited transmon level
  introduces additional dispersive shifts that could affect the multitone selectivity.
\end{enumerate}

%% ============================================================
\section{Reproducibility Summary}
%% ============================================================

\begin{center}
\begin{tabular}{lll}
\toprule
Study & Rerun status & Conclusion \\
\midrule
corrected\_sqr metric & Data read, Study~4 rerun confirms key case & \passmark Consistent \\
parameterized waveform & Study~4 rerun reveals failure at $\nact = 4$ & \passmark Consistent \\
ideal\_sqr direct vs echoed & Study~4 rerun confirms 0.824 joint fidelity & \passmark Consistent \\
native\_multitone validation & Full rerun 2026-03-28, results match exactly & \passmark Reproducible \\
multitone arbitrary fock & Data consistent with definition & \passmark Consistent \\
strong validation & Data consistent; best strict 0.9957 confirmed & \passmark Consistent \\
native rich multitone & Data read, key metrics confirmed from JSON & \passmark Consistent \\
the definitive study & Full rebuild rerun 2026-03-28 (37\,s), all checks pass & \passmark Reproducible \\
cluster state unified & Data read, ground-sector summary confirmed & \passmark Consistent \\
\bottomrule
\end{tabular}
\end{center}

\noindent
All nine studies are internally consistent. The two studies where full reruns
were performed (\texttt{native\_multitone\_sqr\_fixed\_multiinput\_validation} and
\texttt{the\_definitive\_ideal\_sqr\_gate\_study}) reproduced exactly.
Python 3.12 with \texttt{cqed-sim 0.1.0} was used for all reruns.

%% ============================================================
\section*{Appendix: Study Provenance and Code Locations}
%% ============================================================
\addcontentsline{toc}{section}{Appendix: Study Provenance and Code Locations}

\begin{center}
\begin{tabular}{lll}
\toprule
Study & Main script & Key data output \\
\midrule
1 & scripts/run\_study.py & data/study\_summary.json \\
2 & scripts/run\_study.py & data/study\_results.json \\
3 & scripts/run\_study.py & data/study\_results.json \\
4 & scripts/run\_study.py & data/study\_summary.json \\
5 & scripts/run\_study.py & data/study\_results.json \\
6 & scripts/run\_study.py & data/study\_results.json \\
7 & scripts/run\_study.py & data/all\_results.json \\
8 & scripts/run\_full\_study.py & data/new\_results/study\_summary.json \\
9 & scripts/run\_design\_space\_study.py & data/ground\_sector\_followup\_summary.json \\
\bottomrule
\end{tabular}
\end{center}

\noindent The cqed\_sim package is installed editably at
\texttt{C:\textbackslash{}Users\textbackslash{}jl82323\textbackslash{}Box\textbackslash{}...cQED\_simulation}.
The current version (0.1.0) introduced a phase-convention correction relative to the
version used in Studies~2--5; this caused the metric discrepancy documented in Section~5.3.
All results quoted in this report use the current patched package.

\end{document}
